---
title: "Homework 5"
subtitle: "Research Methods, Spring 2024"
author: "Taeyoung Yim"
format:
  pdf:
    output-file: "yim-t-hwk5-3"
    output-ext:  "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra, stringr, data.table, gdata, acs, tidyr)
```

```{r}
#| include: false
#| eval: true

data <- read_tsv('data/input/acs_medicaid.txt')

data <- data %>%
  mutate(perc_private = (ins_employer + ins_direct)/adult_pop,
         perc_public = (ins_medicare + ins_medicaid)/adult_pop,
         perc_ins = (adult_pop - uninsured)/adult_pop,
         perc_unins = uninsured/adult_pop,
         perc_employer = ins_employer/adult_pop,
         perc_medicaid = ins_medicaid/adult_pop,
         perc_medicare = ins_medicare/adult_pop,
         perc_direct = ins_direct/adult_pop) %>%
  filter(! State %in% c("Puerto Rico", "District of Columbia"))
```

\newpage
# Summarize the Data
\vspace{.2in}
\noindent 1. Plot the share of the adult population with direct purchase health insurance over time.<br>

```{r, echo = FALSE}
#| label: fig-directplot
#| fig-cap: "Share of Individuals with Direct Purchase"
direct.plot <- data %>% group_by(year) %>% summarize(mean=mean(perc_direct)) %>%
  ggplot(aes(x=year,y=mean)) + geom_line() + geom_point() + theme_bw() +
  labs(
    x="Year",
    y="Fraction with Direct Purchase",
    title="Share of Direct Purchase Insurance over Time"
  ) +
  geom_vline(xintercept=2013.5, color="green")
direct.plot
```

\newpage
\vspace{.2in}
\noindent 2. Discuss the reduction in direct purchase health insurance in later years. Can you list a couple of policies that might have affected the success of the direct purchase insurance market?<br>

Based on what I found through researching, two policies that affected the success of the direct puchase insurance market are all related to ACA and they are the following: (1) denial of funding marketing, which lower the public awareness about ACA, and (2) promotion of AHPs and HRAs, which offered more affordable plans, affecting the direct purchase in health insurance. 

\newpage
\vspace{.2in}
\noindent 3.Plot the share of the adult population with Medicaid over time.<br>
```{r,  echo = FALSE}
#| label: fig-medicaidplot
#| fig-cap: "Share of Individuals with Medicaid"
medicaid.plot <- data %>% group_by(year) %>% summarize(mean=mean(perc_medicaid)) %>%
  ggplot(aes(x=year,y=mean)) + geom_line() + geom_point() + theme_bw() +
  labs(
    x="Year",
    y="Fraction with Medicaid",
    title="Share of Medicaid Insurance over Time"
  ) +
  geom_vline(xintercept=2013.5, color="green")
medicaid.plot
```


\newpage
\vspace{.2in}
\noindent 4.Plot the share of uninsured over time, separately by states that expanded Medicaid in 2014 versus those that did not. Drop all states that expanded after 2014.<br>
```{r,  echo = FALSE}
#| label: fig-uninsuranceplot
#| fig-cap: "Average Uninsured by Medicaid Expansion"
library(tidyverse)  

ins.plot.dat <- data %>% filter(expand_year==2014 | is.na(expand_year), !is.na(expand_ever)) %>%
  mutate(perc_unins=uninsured/adult_pop) %>%
  group_by(expand_ever, year) %>% summarize(mean=mean(perc_unins))

ins.plot <- ggplot(data=ins.plot.dat, aes(x=year,y=mean,group=expand_ever,linetype=expand_ever)) + 
  geom_line() + geom_point() + theme_bw() +
  geom_vline(xintercept=2013.5, color="green") +
  geom_text(data = ins.plot.dat %>% filter(year == 2016), 
            aes(label = c("Non-expansion","Expansion"),
                x = year + 1,
                y = mean)) +
  guides(linetype="none") +
  labs(
    x="Year",
    y="Fraction Uninsured",
    title="Share of Uninsured over Time"
  )
ins.plot
```

\newpage
# Estimate ATEs
For the rest of the assignment, we’re going to apply the difference-in-differences estimator to the question of Medicaid expansion and uninsurance.
\vspace{.2in}
\noindent 5.Calculate the average percent of uninsured individuals in 2012 and 2015, separately for expansion and non-expansion states. Present your results in a basic 2x2 DD table.<br>
```{r, echo = FALSE}
#| tbl-cap: "DD Table for Medicaid Expansion"
dd.table <- data %>% 
  filter(is.na(expand_year) | expand_year==2014) %>%
  filter(year %in% c(2012, 2015)) %>%  
  group_by(expand_ever, year) %>%
  summarize(uninsured=mean(perc_unins))

dd.table <- pivot_wider(dd.table, names_from="year", names_prefix="year", values_from="uninsured") %>% 
  ungroup() %>%
  mutate(expand_ever=case_when(
    expand_ever==FALSE ~ 'Non-expansion',
    expand_ever==TRUE ~ 'Expansion')
  ) %>%
  rename(Group=expand_ever,
         Pre=year2012,
         Post=year2015)
knitr::kable(dd.table, 
             col.names=c("Group", "Pre", "Post"),
             digits=2,
             booktabs = TRUE,
             escape=F,
             align="l",
             format='latex') %>% 
  kable_styling(full_width=F,
                latex_options="hold_position")
```

\newpage

\vspace{.2in}
\noindent 6.Estimate the effect of Medicaid expansion on the uninsurance rate using a standard DD regression estimator, again focusing only on states that expanded in 2014 versus those that never expanded.<br>

```{r, echo=FALSE}
#| label: tbl-ddregs
#| tbl-cap: "DD Estimates for Medicaid Expansion"
reg.data <- data %>% mutate(post=(year>=2014),
                                  treat=post*expand_ever) %>%
  filter(is.na(expand_year) | expand_year==2014)

dd.est <- lm(perc_unins~post + expand_ever + treat, data=reg.data)
fe.est <- feols(perc_unins~treat | State + year, data=reg.data)

summary(dd.est)
```

\newpage
\vspace{.2in}
\noindent 7.Include state and year fixed effects in your estimates. Try using the lfe or fixest package to estimate this instead of directly including the fixed effects.<br>

```{r, echo=FALSE}
#| label: tbl-ddregs2
#| tbl-cap: "DD Estimates for Medicaid Expansion with TWFE"
reg.data <- data %>% mutate(post=(year>=2014),
                                  treat=post*expand_ever) %>%
  filter(is.na(expand_year) | expand_year==2014)

dd.est <- lm(perc_unins~post + expand_ever + treat, data=reg.data)
fe.est <- feols(perc_unins~treat | State + year, data=reg.data)

summary(fe.est)
```
\newpage
\vspace{.2in}
\noindent 8.Repeat the analysis in question 7 but include all states (even those that expanded after 2014). Are your results different? If so, why?<br>
```{r,  echo = FALSE}
#| label: tbl-ddregs3
#| tbl-cap: "DD Estimates for Medicaid Expansion with Staggered Treatment"
reg.data2 <- data %>% 
  mutate(treat=case_when(
    year>=expand_year & !is.na(expand_year) ~ 1,
    is.na(expand_year) ~ 0,
    year<expand_year & !is.na(expand_year) ~ 0)
  )
fe.est2 <- feols(perc_unins~treat | State + year, data=reg.data2)
summary(fe.est2)

```
The result is a slightly different since the fixed effect that is added is giving more accuracy to the data analysis. Also, having more observation can also affect the result slightly.  

\newpage
\vspace{.2in}
\noindent 9.Provide an “event study” graph showing the effects of Medicaid expansion in each year. Use the specification that includes state and year fixed effects, limited to states that expanded in 2014 or never expanded.<br>
```{r,  echo = FALSE}
#| label: fig-eventstudy1
#| fig-cap: "Event Study with Common Treatment Time"
mod.twfe <- feols(perc_unins~i(year, expand_ever, ref=2013) | State + year,
                  cluster=~State,
                  data=reg.data)
iplot(mod.twfe, 
      xlab = 'Time to treatment',
      main = '')
```

\newpage
\vspace{.2in}
\noindent 10.Repeat part 9 but again include states that expanded after 2014. Note: this is tricky…you need to put all states onto “event time” to create this graph.<br>

```{r,  echo = FALSE}
#| label: fig-eventstudy2
#| fig-cap: "Event Study with Staggered Treatment"
reg.data2 <- reg.data2 %>%
  mutate(time_to_treat=ifelse(expand_ever==TRUE, year-expand_year, -1),
         time_to_treat=ifelse(time_to_treat<=-4, -4, time_to_treat))

mod.twfe2 <- feols(perc_unins~i(time_to_treat, expand_ever, ref=-1) | State + year,
                  cluster=~State,
                  data=reg.data2)
iplot(mod.twfe2, 
      xlab = 'Time to treatment',
      main = '')
```